Improving Software Maintainability through Automated Refactoring of Code Clones

### 2019-10-25 saner/paper.tex
Updates from Overleaf

\subsubsection{Preconditions}
Clones detected by current detection techniques \cite{kamiya2002ccfinder, deissenboeck2010flexible, jiang2007deckard, cordy2011nicad} are subject to a set of preconditions to determine whether they can be refactored \cite{tsantalis2015assessing}:
\begin{enumerate}
  \item The parameterization of differences between the matched statements should not break existing data-, anti-, and output-dependences.
  \item The unmatched statements should be movable before or after the matched statements without breaking existing data-, anti-, and output-dependences.
  \item The duplicated code fragments should return at most one variable of the same type.
  \item Matched branching (\texttt{break}, \texttt{continue}) statements should be accompanied with corresponding matched loop statements.
  \item Matched variables having different subclass types should call only methods that are declared in the common superclass or are being overridden in the respective subclasses.
  \item The parameterization of fields belonging to differences between the mapped statements is possible only if they are not modified.
  \item The parameterization of method calls belonging to differences between the mapped statements is possible only if they do not return a void type.
  \item The mapped statements within the clone fragments should not contain any conditional return statements.
\end{enumerate}

@@

\subsection{Clone context analysis}
Golomingi~\cite{koni2001scenario} explores mapping the relation between clone instances to refactoring methods. The author analyses the refactoring methods described by Martin Fowler \cite{fowler1999refactoring} and analyzes what refactoring methods can be used to refactor clones with what inheritance relations. The identified clone relations are: Ancestor, Common Hierarchy, First Cousin, Same Method, Sibling, Single Class, Superclass and Unrelated. We extend this list with several more fine-grained relations, suitable for automatic refactoring (see Sec.~\ref{sec:relation}).

@@

Apart from these criteria, we found that the output of these clone detection tools cannot be post-processed to find the clone types proposed in Sec.~\ref{sec:rtypes}. This is mainly because the clones of T2R are not a subset of type 2 clones and will thus require an analysis of the entire system. % TODO

@@

\subsection{Clone detection}\label{sec:detection}
We use an AST-based method to detect clones. Clones are detected on a statement level: only full statements are considered as clones. We limit the variability between indicated expressions (see Sec.~\ref{sec:t2r}) by a threshold. This threshold is the percentage of different expressions against the total number of expressions in the source code:
\begin{equation}\label{eq:type2r}
\text{Variability}=\frac{\text{Different expressions}}{\text{Total expressions}}*100
\end{equation}

After all clones have been detected, CR determines whether clone classes can be merged into gapped clones (see Sec.~\ref{sec:t3r}). The maximum size of the gap is limited by a threshold. This threshold is the percentage of (not-cloned) statements in the gap against the sum of statements in both clones surrounding it. Unlike the expression variability threshold, this threshold can exceed 100\%:
\begin{equation}\label{eq:type3r}
\text{Gap Size}=\frac{\text{Statements in gap}}{\text{Statements in clones}}*100
\end{equation}

To verify the correctness of all detected clones, we ran the tool over a large project and manually checked the output. We also created a set of control projects to test the correctness for many edge cases.

@@

\subsection{Context mapping} \label{sec:context}
After clones are detected, we map the context of these clones. We have identified two properties of clones as their context: relation and location. We identify categories for each of these properties, to get a detailed insight into the context of clones.

\subsubsection{Relation}\label{sec:relation}
Clone instances in a clone class may have a relation with each other through inheritance. This relation has a big impact on how the clone should be refactored~\cite{fontana2015duplicated}. We define the following categories to map the relation between clone instances in a clone class. These categories do not map external classes (classes outside the project, e.g. belonging to a library) unless explicitly stated:
\begin{itemize}
    \item \textbf{Common Class}: All clone instances are in the same class.
    \begin{itemize}
        \item \textbf{Same Method}: All clone instances are in the same method.
        \item \textbf{Same Class}: All clone instances are in the same class.
    \end{itemize}
    \item \textbf{Common Hierarchy}: All clone instances are in the same inheritance hierarchy.
    \begin{itemize}
        \item \textbf{Superclass}: Clone instances reside in a class or its parent class.
        \item \textbf{Sibling Class}: All clone instances reside in classes with the same parent class.
        \item \textbf{Ancestor Class}: All clone instances reside in a class, or any of its recursive parents.
        \item \textbf{First Cousin}: All clone instances reside in classes with the same grandparent class.
        \item \textbf{Same Hierarchy}: All clone instances are part of the same inheritance hierarchy.
    \end{itemize}
    \item \textbf{Common Interface}: All clone instances are in classes that have the same interface.
    \begin{itemize}
        \item \textbf{Same Direct Interface}: All clone instances are in classes that have the same interface.
        \item \textbf{Same Class}: All clone instances are in an inheritance hierarchy that contains the same interface.
    \end{itemize}
    \item \textbf{Unrelated}: All clone instances are in classes that have the same interface.
    \begin{itemize}
        \item \textbf{No Direct Superclass}: All clone instances are in classes that have the Object class as their parent.
        \item \textbf{No Indirect Superclass}: All clone instances are in a hierarchy that contains a class that has the Object class as their parent.
        \item \textbf{External Superclass}: All clone instances are in classes the same external class as their parent.
        \item \textbf{Indirect External Ancestor}: All clone instances are in a hierarchy that contains a class that has an external class as their parent.
    \end{itemize}
\end{itemize}

@@

\subsubsection{Location}\label{sec:location}
The location of a clone instance determines what refactoring techniques can be applied to refactor such clones. We define the following categories by which we analyze the location of clones:
\begin{enumerate}
  \item \textbf{Full Method/Constructor/Class/Interface/Enumeration:} A clone that spans a full class, method, constructor, interface or enumeration, including its declaration.
  \item \textbf{Partial Method/Constructor Body:} A clone that spans (a part of) the body of a method/constructor.
  \item \textbf{Several Methods:} A clone that spans over two or more methods, either fully or partially, but does not span anything but methods.
  \item \textbf{Only Fields:} A clone that spans only global variables.
  \item \textbf{Other:} Anything that does not match with above-stated categories.
\end{enumerate}

@@

\subsection{Tool Validation}
We have validated the correctness of CR through unit tests and empirical validation. First, we created a set of 57 control projects to verify the correctness in many (edge) cases. These projects contain clones for each identified relation, location, and refactorability category to see whether they get correctly identified. Next, we run the tool over the corpus and manually verify samples of the acquired results. This way, we check both the correctness of the identified clones, their context, and their proposed refactoring.

>>>>

\subsubsection{Preconditions}
Clones detected by current detection techniques \cite{kamiya2002ccfinder, deissenboeck2010flexible, jiang2007deckard, cordy2011nicad} are subject to a set of preconditions to determine whether they can be refactored \cite{tsantalis2015assessing}:
\begin{enumerate}
  \item The parameterization of differences between the matched statements should not break existing data-, anti-, and output-dependencies.
  \item The unmatched statements should be movable before or after the matched statements without breaking existing data-, anti-, and output-dependencies.
  \item The duplicated code fragments should return at most one variable of the same type.
  \item Matched branching (\texttt{break}, \texttt{continue}) statements should be accompanied with corresponding matched loop statements.
  \item Matched variables having different subclass types should call only methods that are declared in the common superclass or are being overridden in the respective subclasses.
  \item The parameterization of fields belonging to differences between the mapped statements is possible only if they are not modified.
  \item The parameterization of method calls belonging to differences between the mapped statements is possible only if they do not return a void type.
  \item The mapped statements within the clone fragments should not contain any conditional return statements.
\end{enumerate}

@@

\subsection{Clone context analysis}\label{sec:rw:contextanalysis}
Golomingi~\cite{koni2001scenario} explores mapping the relation between clone instances to refactoring methods. The author analyses the refactoring methods described by Martin Fowler \cite{fowler1999refactoring} and analyzes what refactoring methods can be used to refactor clones with what inheritance relations. The identified clone relations are: Ancestor, Common Hierarchy, First Cousin, Same Method, Sibling, Single Class, Superclass and Unrelated. We extend this list with several more fine-grained relations, suitable for automatic refactoring (see Sec.~\ref{sec:relation}).

@@

Apart from these criteria, we found that the output of these clone detection tools cannot be post-processed to find the clone definition proposed in Sec.~\ref{sec:research}. This is mainly because these clones are not necessarily a subset of clones detected by these tools and will thus require an analysis of the entire system.

@@

\subsection{Clone detection}\label{sec:detection}
We use an AST-based method to detect clones. Clones are detected on a statement level: only full statements are considered as clones. In Sec.~\ref{sec:t2r} we described that we limit the variability between variable, literal and method reference expressions by a threshold. This threshold is the percentage of different expressions against the total number of expressions in the source code:
\begin{equation}\label{eq:type2r}
\text{Variability}=\frac{\text{Different expressions}}{\text{Total expressions}}*100
\end{equation}

After all clones have been detected, CR determines whether clone classes can be merged into gapped clones (see Sec.~\ref{sec:t3r}). The maximum size of the gap is limited by a threshold. This threshold is the percentage of (not-cloned) statements in the gap against the sum of statements in both clones surrounding it:
\begin{equation}\label{eq:type3r}
\text{Gap Size}=\frac{\text{Statements in gap}}{\text{Statements in clones}}*100
\end{equation}

Note that unlike the expression variability threshold, this threshold can exceed 100\%. This is because, in theory, the gap can be larger than the clones surrounding it.

% To verify the correctness of all detected clones, we ran the tool over a large project and manually checked the output. We also created a set of control projects to test the correctness for many edge cases.

@@

\subsection{Context mapping} \label{sec:context}
After clones are detected, we map the context of these clones. We identify two properties of clones as their context: relation~\cite{fontana2012duplicated} and location~\cite{tairas2011representing}. We identify categories for each of these properties, to get a detailed insight into the context of clones.

\subsubsection{Relation}\label{sec:relation}
Clone instances in a clone class may have a relation with each other through inheritance. This relation has a big impact on how the clone should be refactored~\cite{fontana2015duplicated}. We define the following categories to map the relation between clone instances in a clone class, partly derived from Fontana et al~\cite{fontana2015duplicated} (see Sec~\ref{sec:rw:contextanalysis}). These categories do not map external classes (classes outside the project, e.g. belonging to a library) unless explicitly stated:
\begin{itemize}
    \item \textbf{Common Class}: All clone instances are in the same class.
    \begin{itemize}
        \item[$\square$] \textbf{Same Method}: All clone instances are in the same method.
        \item \textbf{Same Class}: All clone instances are in the same class.
    \end{itemize}
    \item \textbf{Common Hierarchy}: All clone instances are in the same inheritance hierarchy.
    \begin{itemize}
        \item \textbf{Superclass}: Clone instances reside in a class or its parent class.
        \item \textbf{Sibling Class}: All clone instances reside in classes with the same parent class.
        \item \textbf{Ancestor Class}: All clone instances reside in a class, or any of its recursive parents.
        \item \textbf{First Cousin}: All clone instances reside in classes with the same grandparent class.
        \item \textbf{Same Hierarchy}: All clone instances are part of the same inheritance hierarchy.
    \end{itemize}
    \item \textbf{Common Interface}: All clone instances are in classes that have the same interface.
    \begin{itemize}
        \item \textbf{Same Direct Interface}: All clone instances are in classes that have the same interface.
        \item \textbf{Same Class}: All clone instances are in an inheritance hierarchy that contains the same interface.
    \end{itemize}
    \item \textbf{Unrelated}: All clone instances are in classes that have the same interface.
    \begin{itemize}
        \item \textbf{No Direct Superclass}: All clone instances are in classes that have the Object class as their parent.
        \item \textbf{No Indirect Superclass}: All clone instances are in a hierarchy that contains a class that has the Object class as their parent.
        \item \textbf{External Superclass}: All clone instances are in classes the same external class as their parent.
        \item \textbf{Indirect External Ancestor}: All clone instances are in a hierarchy that contains a class that has an external class as their parent.
    \end{itemize}
\end{itemize}

@@

\subsubsection{Location}\label{sec:location}
The location of a clone instance helps determine what refactoring techniques can be applied to refactor such clones. We define the following categories by which we analyze the location of clones:
\begin{enumerate}
  \item \textbf{Full Method/Constructor/Class/Interface/Enumeration:} A clone that spans a full class, method, constructor, interface or enumeration, including its declaration.
  \item \textbf{Partial Method/Constructor Body:} A clone that spans (a part of) the body of a method/constructor.
  \item \textbf{Several Methods:} A clone that spans over two or more methods, either fully or partially, but does not span anything but methods.
  \item \textbf{Only Fields:} A clone that spans only global variables.
  \item \textbf{Other:} Anything that does not match with above-stated categories.
\end{enumerate}

@@

\subsection{Tool Evaluation}
We assess the correctness of CR through unit tests and empirical validation. First, we create a set of 57 control projects to verify the correctness in many (edge) cases. These projects contain clones for each identified relation, location, and refactorability category to see whether they get correctly identified. Next, we run the tool over the corpus and manually verify samples of the acquired results. This way, we check both the correctness of the identified clones, their context, and their proposed refactoring.
